---
title: "Coursera Reproducible Research Project1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the libraries and download and preprocess data:

```{r }
library(ggplot2)
library(dplyr)
temp=tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp, mode="wb")
unzip(temp, "activity.csv")
activity <- read.csv("activity.csv")
```
###Get data without NAs
```{r }
newactivity <- na.omit(activity) 
```
## Mean total number of steps taken per day omitting missing values
```{r }
totalsteps=newactivity %>%
  group_by(date) %>%
  summarize(sumsteps = sum(steps, na.rm = TRUE))
```
###Histogram of number of steps taken per day
```{r }
hist(totalsteps$sumsteps,xlab="Totalsteps",ylab="Frequency",main ="Toal number of steps taken each day" )
```

## Mean of the total number of steps taken per day
```{r }
meansteps= mean(totalsteps$sumsteps, na.rm = TRUE)
meansteps

```
## Median of the total number of steps taken per day
```{r }
mediansteps=median(totalsteps$sumsteps, na.rm = TRUE)
mediansteps
```
##Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r }
avgsteps=activity %>%
  group_by(interval) %>%
  summarize(mean = mean(steps, na.rm = TRUE))
plot(avgsteps$interval,avgsteps$mean,type="l",xlab="Interval",ylab="Average Steps",main ="5 minutes interval vs average steps taken")
```

##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r }
avgsteps[avgsteps$mean==max(avgsteps$mean),]
```
The 5-minute interval 835 contains the maximum number of steps(206 steps).

## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r }
naactivity=activity[is.na(activity),]
nrow(naactivity)
```
Total number of rows with missing values is 2304.

##Impute missing values using mean of that interval and create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r }
##  Get the mean for each 5 mins interval
meaninterval=activity %>%
  group_by(interval) %>%
  summarize(mean = mean(steps, na.rm = TRUE))
## Merge the main dataset with the dataset containing mean of each 5 minutes interval
total <- merge(activity,meaninterval,by="interval") 
## Assign the 5 minutes interval mean value to the NA steps columns
total$steps <- ifelse(is.na(total$steps) == TRUE, total$mean, total$steps) 

```
## Mean Total no of steps taken each day
```{r }
totalimpute=total %>%
  group_by(date) %>%
  summarize(sumsteps = sum(steps, na.rm = TRUE))

hist(totalimpute$sumsteps,xlab="Totalsteps",ylab="Frequency",main ="Toal number of steps taken each day" )
```

## Mean of the total number of steps taken per day
```{r }
mean=mean(totalimpute$sumsteps)
mean
```
## Median of the total number of steps taken per day
```{r }
  median=median(totalimpute$sumsteps)
median
```

## Differences in activity patterns in weekdays and weekend
```{r }
total$day <- weekdays(as.Date(total$date))

total$day1 <- ifelse((total$day == 'Saturday' | total$day == 'Sunday' ), 'weekend', 'weekday')  

total_weekday=total[total$day1=='weekday',]
total_weekend=total[total$day1=='weekend',]

avgsteps_weekday=total_weekday %>%
  group_by(interval) %>%
  summarize(mean = mean(steps, na.rm = TRUE))

avgsteps_weekend=total_weekend %>%
  group_by(interval) %>%
  summarize(mean = mean(steps, na.rm = TRUE))
# Add a column day to indicate if its weekday or weekend
avgsteps_weekday$day='weekday'
avgsteps_weekend$day='weekend'
total_week=rbind(avgsteps_weekday,avgsteps_weekend)

p<- ggplot(total_week, aes(x =interval , y=mean, color=day)) +
       geom_line() +
       labs(title = "5 minutes interval vs average steps taken", x = "Interval", y = "Steps") +
       facet_wrap(~day, ncol = 1, nrow=2)
p

```